Sparse Matrix Vector Multiplication with Multiple vectors and Multiple Matrices on the MIC Architecture \todo{? more application oriented''}

### 2013-10-19 ieee_paper.tex
minor intro

In this paper, we consider the case of local stencils with the number
of nodes at each point and encode the sparse matrix in ELLPACK
format. We increase the number of operations relative to memory
bandwidth by calculating four derivatives of four different functions,
or 16 different derivatives. We demonstrate a novel implementation on
the MIC architecture, taking into account its advanced swizzling and
channel interchange features. We present benchmarks that show an
almost order of magnitude increase in speed compared to efficient
implementations of a single derivative. We explain the results through
consideration of operation count versus memory bandwidth.
\end{abstract}

\begin{IEEEkeywords}
OpenMP; MIC; spmv; sparse matrix; Radial Basis Function;
\end{IEEEkeywords}

@@

\section{Introduction}
The multiplication of sparse matrix by a dense vector (SpMV) is an
important kernel in many applied fields such as fluid dynamics 
\cite{journals/tog/BolzFGS03} 
to graph analysis (recommender systems~\cite{Brin98}, graph
drawing~\cite{Koren05}). So improving the performance of SpMV has
captured the interest of many researcher; including the development of
various implementations for multi-core
CPUs~\cite{Buluc2009SPAA,Williams07} and GPUs~\cite{Bell08,
  conf/ipps/KreutzerHWFBB12,
  journals/concurrency/VazquezFG11,kumar2012accelerating}. The main
challenge that is faced to obtain good performance for SpMV is that
the operation are conducted using memory location that are irregular
and often unpredictible. That makes the kernels mostly memory
bound and there is a significant instruction overhead per flop.

>>>>

In this paper, we consider the case of local stencils with the number
of nodes at each point and encode the sparse matrix in ELLPACK
format. We increase the number of operations relative to memory
bandwidth by calculating four derivatives of four different functions,
or 16 different derivatives. We demonstrate a novel implementation on
the MIC architecture, taking into account its advanced swizzling and
channel interchange features. We present benchmarks on a real data
set that show an almost eightfold increase in speed compared to
efficient implementations of a single derivative, reaching a
performance of almost 140~Gflop/s. We explain the results through
consideration of operation count versus memory bandwidth.
\end{abstract}

\begin{IEEEkeywords}
MIC; SIMD; SpMV; SpMM; Sparse Matrix; Radial Basis Function;
\end{IEEEkeywords}

@@

\section{Introduction}
The multiplication of sparse matrix by a dense vector (SpMV) is an
important kernel in many applied fields such as fluid dynamics 
\cite{journals/tog/BolzFGS03}, recommendation systems~\cite{Brin98} and graph
drawing~\cite{Koren05}). Naturally, improving the performance of SpMV has
captured the interest of many researchers; including the development of
various implementations for 
CPUs~\cite{Buluc2009SPAA,Williams07} and GPUs~\cite{Bell08,
  conf/ipps/KreutzerHWFBB12,
  journals/concurrency/VazquezFG11,kumar2012accelerating}. The main
challenge  to obtain good performance for SpMV is that
the operations are conducted using memory location that are irregular. This makes the kernels mostly memory
bound and there is a significant instruction overhead per flop.
